Abstractions and exemplars: the measure noun phrase alternation in German

### 2018-01-01 Latex/article.tex
Done with everything except conclusion.

\begin{acknowledgement}
  I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, Ulrike Sayatz, and Christian Zimmer for valuable discussions and comments.
  Also, I would like to thank the reviewers for Cognitive Linguistics as well as associate editor Dagmar Divjak for insightful comments which helped to improve the quality of this paper significantly.
  Furthermore, I thank Ulrike Sayatz for helping me to recruit the participants for the experiments.
  Elizabeth Pankratz thankfully also fixed my English.
  Finally, I am grateful to my student assistants Kim Maser, for her work on the annotation of the concordances, and Luise Rißmann, for supervising most of the experiments.
  The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).
\end{acknowledgement}

>>>>

\begin{acknowledgement}
  I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, Ulrike Sayatz, and Christian Zimmer for valuable discussions and comments.
  Also, I would like to thank the reviewers for Cognitive Linguistics as well as associate editor Dagmar Divjak for insightful comments which helped to improve the quality of this paper significantly.
  Furthermore, I thank Ulrike Sayatz for helping me to recruit the participants for the experiments.
  Elizabeth Pankratz thankfully also fixed my English.
  Finally, I am grateful to my student assistants Kim Maser for her work on the annotation of the concordances and Luise Rißmann for supervising most of the experiments.
  The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).
\end{acknowledgement}

### 2018-01-01 Latex/sections/3corpus.tex
Done with everything except conclusion.

@@

For this pre-study, each of the 2,365 combinations of measure noun lemma and kind noun lemma were queried in the \NACb\ and the \PGCd.
Each of the resulting 2,365 concordances was scaled down randomly to a size of maximally 100 sentences, and from the resulting 35,766 sentences, 5,000 were randomly sampled for the statistical analysis.
All features described in Section~\ref{sec:variablesandannotation} were annotated except for the ones which do not apply in the non-alternating case (\textit{Kindattraction}, \textit{Measureattraction}, and \textit{Measurecase}).
The aggregated counts for each concordance were analysed in the form of a multilevel logistic regression model using R \citep{R} and the \textit{lme4} package \citep{Bates2010,BatesEa2015}.%
\footnote{An alternative \textit{BOBYQA} optimiser from the \textit{nloptr} package \citep{Johnson2017} was used for all fits with \textit{lme4} reported in this paper.}

@@

The coefficient estimates are specified in Table~\ref{tab:smalltable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as characterising a configuration typical of the \PGCd.
For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with 1,000 replications and using the percentile method for the calculation of the intervals.
The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:smalltable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
The column \textit{CI excludes 0} shows an asterisk for those intervals that do not include 0.
Furthermore, for each regressor, a p-value was obtained by dropping the regressor from the full model, re-estimating the nested model, and comparing it to the full model.
Instead of inexact Wald approximations and Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
I call the corresponding value $p_{\text{PB}}$, and it is given in the respective columns in Table~\ref{tab:smalltable}.
Regressors which did not reach sig=0.05 in in the PBmodcomp test (\textit{Kindfreq}, \textit{Kindgender}, and \textit{Measurefreq}) were removed from the model, appear in brackets in Table~\ref{tab:smalltable}, and have no coefficient estimates.
The overall intercept is -5.370 and reoresents \textit{Cardinal=Yes}, \textit{Measureclass=Physical} and 0 for all (z-transformed) numeric regressors.
The standard deviation in the random intercepts is 1.157 for \textit{Measurelemma} ($p_{\text{PB}}=0.001$) and 0.923 for \textit{Kindlemma} ($p_{\text{PB}}=0.0.002$).
Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.278$ and $R^2_c=0.566$ (see \citealp{Gries2015} for a basic introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
This means that the effect of the fixed effects (marginal $R^2$) is weak to intermediate, and that lemma-specific effects (conditional minus marginal $R^2$) are equally strong with a difference between the two coefficients of 0.289.

This show that the main prototypical factors \textit{Cardinal} and \textit{Measureclass} indeed play the predicted role
The abstract non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACb.
At the other end of the scale, nouns denoting natural portions like \textit{Haufen} `heap', \textit{Bündel} `bundle', \textit{Schluck} `gulp' favour the \PGCd.
Also, The presence of a cardinal modifier clearly favours the \NACb.
The stylistic factor \textit{Genitives} is confirmed to have an influence as predicted inasmuch as a general lack of genitives (which is encoded as a \textit{higher} value of the variable) favours the \NACb.
\textit{Badness} points into the wrong direction, but it barely reaches sig=0.05, and its confidence interval includes 0.
The lemma frequency regressors both clearly fail the PBmodcomp test.
I will return to these results in Section~\ref{sec:interpretation}.

@@

@@

Using signed logarithmised collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) (with smoothing to avoid log(0) operations) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
While the attraction measures reach satisfying sig levels in the PBmodcomp test (0.020 for \textit{Kindattraction} and 0.001 for \textit{Measureattraction}), the $p_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.
The coefficients of determination drop to $R_m^2=0.135$ and $R^2_c=0.636$.

\subsection{Interpretation}
\label{sec:interpretation}

%\hl{As collostructional analysis is relatively popular in some cognitively oriented areas of corpus linguistics, I tried using collexeme strength as a regressor (with smoothing to remedy the mathematical problems), and the results were unsatisfactory compared to the simple quotient used here.
%The corresponding regressor did not even reach sig=0.1 in the PBmodcomp test.
%Both the cognitive underpinnings and the math behind collostructional analysis are still much debated (see prominently \citealp{SchmidKuechenhoff2013,KuechenhoffSchmid2015}), I see this result }

% ATTRACTION

@@

An interesting picture emerges for the lemma frequencies.
A higher-than-average lemma frequency of measure nouns favours the \NACa\ ($\beta_{\text{Measurefreq}}=-0.231$, $\mpPB=0.005$), which is as expected if we assume at least a tendency for highly grammaticalised items to be more frequent.
With kind nouns, higher frequency seems to favour the \PGCa ($\beta_{\text{Kindfreq}}=0.146$, $\mpPB=0.095$).
However, there is no clear theoretical interpretation (see Section~\ref{sec:analyses}), and the estimate is imprecise (not significant at $\alpha=0.05$, see above).
The effect can therefore be ignored or treated as a nuisance variable.

% GRAMMATICALISATION
% => lemmas and lemma classes

@@

In Section~\ref{sec:analyses}, it was also hypothesised that classes of measure nouns with a higher degree of grammaticalisation should favour the \NACa.
The \textit{Measureclass} second-level predictor was successfully estimated ($\mpPB=0.001$).
Looking at the effect plot in Figure~\ref{fig:eff:measureattraction}, it is evident that abstract non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACa.
At the other end of the scale, nouns denoting natural portions like \textit{Haufen} `heap', \textit{Bündel} `bundle', \textit{Schluck} `gulp' favour the \PGCa.
These are referential nouns, confirming the hypothesis that it is prototypical of the PGC to contain two referential nouns, while the NAC prototypically only contains one (the kind noun).

% CARDINALS

@@

I now turn to the predicted effect of cardinals as modifiers of the measure noun.
Figure~\ref{fig:eff:leftcontext} shows that cardinals indeed influence the choice of the alternant ($\mpPB=0.001$), and that cardinals have a strong tendency to co-occur with the \NACa.
This effect was predicted in Section~\ref{sec:analyses}.

% REGISTER
% => Badness
% => Genitives

The style-related proxy variables point in the expected direction.
Increased \textit{Badness} of the document favours the \NACa\ ($\beta_{\text{Badness}}=-0.152$, $\mpPB=0.002$), and so does a lower density of genitives ($\beta=-0.693$, $\mpPB=0.001$).
While these are merely proxies to style (and partially circular in the case of \textit{Genitives}), this result can at least encourage future work into stylistic effects.

% NUISANCE
% => dative effect
% => frequency

The influence of \textit{Measurecase} ($\mpPB=0.001$) is as predicted in previous analyses (see Section~\ref{sec:analyses}).
A measure noun in the dative favours the \PGCa\ with $\beta_{\text{MeasurecaseDat}}=0.705$ (compared to the nominative, which is on the intercept).
Although \textit{Measurecase} is a nuisance variable in the context of this study, convergence with previous work strengthens its validity.

>>>>

\subsubsection{Sampling and annotation}

@@

For this pre-study, each of the 2,365 combinations of measure noun lemma and kind noun lemma were queried in the \NACb\ and the \PGCd.
Each of the resulting 2,365 concordances was scaled down randomly to a size of maximally 100 sentences, and from the resulting 35,766 sentences, 5,000 were randomly sampled for the statistical analysis.
All features described in Section~\ref{sec:variablesandannotation} were annotated except for the ones which do not apply in the non-alternating case (\textit{Kindattraction}, \textit{Measureattraction}, and \textit{Measurecase}).

\subsubsection{Model}

The annotated concordance was analysed in the form of a multilevel logistic regression model using R \citep{R} and the \textit{lme4} package \citep{Bates2010,BatesEa2015}.%
\footnote{An alternative \textit{BOBYQA} optimiser from the \textit{nloptr} package \citep{Johnson2017} was used for all fits with \textit{lme4} reported in this paper.}

@@

The coefficient estimates are specified in Table~\ref{tab:smalltable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as characterising a configuration typical of the \PGCd.
For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with 1,000 replications and using the percentile method for the calculation of the intervals.
The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:smalltable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
The column \textit{CI excludes 0} shows an asterisk for those intervals that do not include 0.
Furthermore, for each regressor, a p-value was obtained by dropping the regressor from the full model, re-estimating the nested model, and comparing it to the full model.
Instead of inexact Wald approximations and Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
I call the corresponding value $p_{\text{PB}}$, and it is given in the respective columns in Table~\ref{tab:smalltable}.
Regressors which did not reach sig=0.05 in in the PBmodcomp test (\textit{Kindfreq}, \textit{Kindgender}, and \textit{Measurefreq}) were removed from the model, appear in brackets in Table~\ref{tab:smalltable}, and have no coefficient estimates.
The overall intercept is -5.370 and reoresents \textit{Cardinal=Yes}, \textit{Measureclass=Physical} and 0 for all (z-transformed) numeric regressors.
The standard deviation in the random intercepts is 1.157 for \textit{Measurelemma} ($p_{\text{PB}}=0.001$) and 0.923 for \textit{Kindlemma} ($p_{\text{PB}}=0.0.002$).
Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.278$ and $R^2_c=0.566$ (see \citealp{Gries2015} for a basic introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).

\subsubsection{Interpretation}

The coefficients of determination indicate that the effect of the fixed effects (marginal $R^2$) is weak to intermediate, and that lemma-specific effects (conditional minus marginal $R^2$) are equally strong with a difference between the two coefficients of 0.289.

The main prototypical factors \textit{Cardinal} and \textit{Measureclass} bot pass the PBmodcomp test at sig=0.05 and play the expected role.
The abstract non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACb.
At the other end of the scale, nouns denoting natural portions like \textit{Haufen} `heap', \textit{Bündel} `bundle', \textit{Schluck} `gulp' favour the \PGCd.
Also, The presence of a cardinal modifier clearly favours the \NACb.
The stylistic factor \textit{Genitives} is confirmed to have an influence as predicted inasmuch as a general lack of genitives (which is encoded as a \textit{higher} value of the variable) favours the \NACb.
\textit{Badness} points into the wrong direction, but it barely reaches sig=0.05, and its confidence interval includes 0.
The lemma frequency regressors both clearly fail the PBmodcomp test.
I will return to these results in Section~\ref{sec:conclusion}.

@@

\subsubsection{Model}

@@

Using signed logarithmised collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) (with smoothing to avoid log(0) operations) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
While the attraction measures reach satisfying sig levels in the PBmodcomp test (0.020 for \textit{Kindattraction} and 0.001 for \textit{Measureattraction}), the $p_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.%
\footnote{My approach to statistical inference is essentially Fisherian \citep{Fisher1935,Fisher1959}, and hence I use the term \textit{sig level} instead of the term \textit{$\alpha$ level} known from null hypothesis significance testing (NHST).
For overviews of systems of statistical inference with special focus on this and related questions, see \citet{Lehmann1993,Lehmann2011} or the shorter \citet{Perezgonzalez2015}.}
The coefficients of determination drop to $R_m^2=0.135$ and $R^2_c=0.636$.

\subsubsection{Interpretation}

@@

An interesting picture emerges for the lemma frequencies.
A higher-than-average lemma frequency of measure nouns favours the \NACa\, which is as expected if we assume at least a tendency for highly grammaticalised items to be more frequent.
With kind nouns, higher frequency seems to favour the \PGCa.
However, there is no clear theoretical interpretation (see Section~\ref{sec:analyses}), and the estimate is imprecise (not significant at sig=0.05).
The effect can therefore be ignored or treated as a nuisance variable.

@@

In Section~\ref{sec:analyses}, it was also hypothesised that classes of measure nouns with a higher degree of grammaticalisation should favour the \NACa.
The \textit{Measureclass} second-level predictor reaches sig=0.05 in the PBmodcomp test.
Looking at the effect plot in Figure~\ref{fig:eff:measureattraction}, it is evident that abstract non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACa.
At the other end of the scale, nouns denoting natural portions like \textit{Haufen} `heap', \textit{Bündel} `bundle', \textit{Schluck} `gulp' favour the \PGCa.
These are referential nouns, confirming the hypothesis that it is prototypical of the PGC to contain two referential nouns, while the NAC prototypically only contains one (the kind noun).

@@

Figure~\ref{fig:eff:leftcontext} shows that cardinals indeed influence the choice of the alternant, and that cardinals have a strong tendency to co-occur with the \NACa.
This effect was predicted in Section~\ref{sec:analyses}.

The style-related proxy variables point in the expected direction.
Increased \textit{Badness} of the document favours the \NACa, and so does a lower density of genitives.
While these are merely proxies to style (and partially circular in the case of \textit{Genitives}), this result can at least encourage future work into stylistic effects.

The influence of \textit{Measurecase} is as predicted in previous analyses (see Section~\ref{sec:analyses}).
A measure noun in the dative favours the \PGCa\ (compared to the nominative, which is on the intercept).
Although \textit{Measurecase} is a nuisance variable in the context of this study, convergence with previous work strengthens its validity.

### 2018-01-01 Latex/sections/4experiments.tex
Done with everything except conclusion.

@@

Then, a multilevel logistic regression was specified with the probability of the \PGCa\ predicted for each sentence by the corpus-based GLMM as the only fixed effect \textit{Modelprediction}.%
\footnote{The document-level variables \textit{Badness} and \textit{Genitives} were set to 0, which is the mean for z-transformed variables.}
A random intercept and slope were added for the individual sentence (item) in order to catch idiosyncrasies of single sentences.
Coefficients were estimated with Maximum Likelihood Estimation (\textit{lmer} function from \textit{lme4}).
The number of observations was \textit{n=}384.

@@

A good amount of the variance can be accounted for by idiosyncrasies of single sentences ($\sigma_{\text{Item}}=1.217$).%
Also, among participants, there are clearly different preferences ($\sigma_{\text{Participant}}=0.412$).
On the extreme ends, one participant chose the \PGCa\ in 13 of 16 cases, and two participants only chose it in 5 of 16 cases.
The regressor \textit{Modelprediction} achieves $\mpPB=0.003$ (1,000 replications) and is estimated at 4.389 relative to an intercept of -1.270.
The confidence interval from a parametric bootstrap (1,000 replications, percentile method) for the regressor is acceptable but slightly large with a lower bound of 1.788 and an upper bound of 6.599.
The pseudo-coefficients of determination are $R^2_{m}=0.185$ and $R^2_{c}=0.455$, which means that roughly 19\% of the variance in the data can be explained by considering only the predictions from the corpus-based GLMM.
This is a weak result for the fixed effects part, but clearly worthy of mention.
The effect display for the single fixed regressor \textit{Modelprediction} is given in Figure \ref{fig:afc:effects}.
The higher the probability of the \PGCa\ predicted from usage data, the more often participants chose the \PGCa\ alternant in the forced-choice task.
A closer look at the results in the form of the spineplot in Figure~\ref{fig:spines} shows, however, that it was likely idiosyncracies in the sentences with a model prediction between 0.5 and 0.6 which spoiled an otherwise much better correlation.
Clearly, a revised and improved experiment might lead to a much better fit in future research.

@@

The reading times were residualised per speaker based on the reading times of all words (not just the targets) by that speaker.
The adjective and the kind noun (\ie\ the constituents bearing the critical case markers) were used as the target region, for instance the bracketed words in the example \textit{zwei Gläser} [\textit{spru\-deln\-des Wasser}] `two glasses of sparkling water'.
Outliers farther than 2 inter-quartile ranges from the mean logarithmised residualised reading time were removed (64 data points), resulting in a total number of \textit{n=}1,152 observations.
An LMM was specified with the logarithmised residual reading times as the response variable.

The probabilities derived from the corpus GLMM (\textit{Modelprediction}) were added as the main regressor of interest.
It should be remembered that the corpus GLMM predicts the probability of the \PGCa.
As a consequence, the higher the GLMM prediction is, the more typical the sentence is for containing the \PGCa.
It is therefore expected that reading times are higher when the value of \textit{Modelprediction} is higher but the sentence contains the \NACa.
However, when the sentence contains the \PGCa, reading times should be lower when \textit{Modelprediction} is higher.
To account for this, an interaction between \textit{Modelprediction} and \textit{Construction} (levels \textit{PGCadj} and \textit{NACadj}) was added to the model.

@@

Table~\ref{tab:exp:spr} shows the coefficient estimates with a 95\% parametric bootstrap confidence interval (1,000 replications, percentile method).
The standard deviation of the participant intercepts is $\sigma_{\text{Participant}}=0.079$ and of the item intercepts $\sigma_{\text{Item}}=0.037$.
Comparing the full model to a model without the main regressor \textit{Modelprediction} (and consequently also without the interaction with \textit{Construction}) in a PB test gives $\mpPB=0.036$.
The pseudo-coefficients of determination are $R^2_m=0.239$ and $R^2_c=0.346$.
There is thus some inter-subject variation and a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities.

>>>>

\subsubsection{Setup, stimuli, and participants}

@@

A multilevel logistic regression was specified with the probability of the \PGCa\ predicted for each sentence by the corpus-based GLMM as the only fixed effect \textit{Modelprediction}.%
\footnote{The document-level variables \textit{Badness} and \textit{Genitives} were set to 0, which is the mean for z-transformed variables.}
A random intercept and slope were added for the individual sentence (item) in order to catch idiosyncrasies of single sentences.
Coefficients were estimated with Maximum Likelihood Estimation (\textit{lmer} function from \textit{lme4}).
The number of observations was \textit{n=}384.

@@

A good amount of the variance can be accounted for by idiosyncrasies of single sentences ($\sigma_{\text{Item}}=1.217$).%
Also, among participants, there are clearly different preferences ($\sigma_{\text{Participant}}=0.412$).
On the extreme ends, one participant chose the \PGCa\ in 13 of 16 cases, and two participants only chose it in 5 of 16 cases.
The regressor \textit{Modelprediction} achieves $\mpPB=0.003$ (1,000 replications) and is estimated at 4.389 relative to an intercept of -1.270.
The confidence interval from a parametric bootstrap (1,000 replications, percentile method) for the regressor is acceptable but slightly large with a lower bound of 1.788 and an upper bound of 6.599.
The pseudo-coefficients of determination are $R^2_{m}=0.185$ and $R^2_{c}=0.455$, which means that roughly 19\% of the variance in the data can be explained by considering only the predictions from the corpus-based GLMM.

\subsubsection{Interpretation}

The coefficients of determination indicate a weak result for the fixed effects part, which is nonetheless worthy of mention (very close to 0.2).
The effect display for the single fixed regressor \textit{Modelprediction} is given in Figure \ref{fig:afc:effects}.
The higher the probability of the \PGCa\ predicted from usage data, the more often participants chose the \PGCa\ alternant in the forced-choice task.
A closer look at the results in the form of the spineplot in Figure~\ref{fig:spines} shows, however, that it was likely idiosyncracies in the sentences with a model prediction between 0.5 and 0.6 which spoiled an otherwise much better correlation.
Clearly, a revised and improved experiment might lead to a much better fit in future research.

@@

The reading times were residualised per speaker based on the reading times of all words (not just the targets) by that speaker.
The adjective and the kind noun (\ie\ the constituents bearing the critical case markers) were used as the target region, for instance the bracketed words in the example \textit{zwei Gläser} [\textit{spru\-deln\-des Wasser}] `two glasses of sparkling water'.
Outliers farther than 2 inter-quartile ranges from the mean logarithmised residualised reading time were removed (64 data points), resulting in a total number of \textit{n=}1,152 observations.

\subsubsection{Model}

An LMM was specified with the logarithmised residual reading times as the response variable.
The probabilities derived from the corpus GLMM (\textit{Modelprediction}) were added as the main regressor of interest.
It should be remembered that the corpus GLMM predicts the probability of the \PGCa.
As a consequence, the higher the GLMM prediction is, the more typical the sentence is for containing the \PGCa.
It is therefore expected that reading times are higher when the value of \textit{Modelprediction} is higher but the sentence contains the \NACa.
However, when the sentence contains the \PGCa, reading times should be lower when \textit{Modelprediction} is higher.
To account for this, an interaction between \textit{Modelprediction} and \textit{Construction} (levels \textit{PGCadj} and \textit{NACadj}) was added to the model.

@@

Table~\ref{tab:exp:spr} shows the coefficient estimates with a 95\% parametric bootstrap confidence interval (1,000 replications, percentile method).
The standard deviation of the participant intercepts is $\sigma_{\text{Participant}}=0.079$ and of the item intercepts $\sigma_{\text{Item}}=0.037$.
Comparing the full model to a model without the main regressor \textit{Modelprediction} (and consequently also without the interaction with \textit{Construction}) in a PB test gives $\mpPB=0.036$.
The pseudo-coefficients of determination are $R^2_m=0.239$ and $R^2_c=0.346$.

An alternative Gaussian generalised additive model with an identity link was also fit (see \citealp{DivjakEa2016}) using the \textit{mgcv} package \citep{Wood2011}.
The full results are included in the data package for this paper, but the fit was not better than with the LMM reported above.
The estimated smoother for the \textit{Modelprediction} variable is essentially linear, and the $R^2$ (corresponding to the marginal $R^2$ of the LMM) was 0.237.

\subsubsection{Interpretation}

The coefficients of determination indicate that there is a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities (marginal $R^2$ and that there some inter-subject variation (conditional minus marginal $R^2$).
